cardinalR: Generating Interesting High-Dimensional Data Structures

### 2025-11-25 paper-cardinalR.Rmd
Try to find why rjtolls have the error

---
title: "cardinalR: Generating Interesting High-Dimensional Data Structures"
abstract: |
  High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association. It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold. Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning. These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets. By controlling the complexity researchers can better understand how different methods work and can be improved. A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets. This enriches the existing toolset of benchmark datasets for evaluating algorithms.
draft: true
author:
  - name: Jayani P. Gamage
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
creative_commons: CC BY
date: "`r Sys.Date()`"
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{float}
output:
  rjtools::rjournal_article:
    keep_md: true
bibliography: paper-cardinalR.bib
editor_options:
  chunk_output_type: inline
---

>>>>

---
title: "cardinalR: Generating Interesting High-Dimensional Data Structures"
abstract: >
  High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association. It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold. Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning. These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets. By controlling the complexity researchers can better understand how different methods work and can be improved. A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets. This enriches the existing toolset of benchmark datasets for evaluating algorithms.
draft: true
author:
  - name: Jayani P. Gamage
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: https://jayanilakshika.netlify.app/
    orcid: 0000-0002-6265-6481
    email: jayani.piyadigamage@monash.edu
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    url: http://www.dicook.org/
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
  - name: Paul Harrison
    affiliation: Monash University
    address: MGBP, BDInstitute, VIC 3800 Australia
    email: paul.harrison@monash.edu
    orcid: 0000-0002-3980-268X
  - name: Michael Lydeamore
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, VIC 3800 Australia
    email: michael.lydeamore@monash.edu
    orcid: 0000-0001-6515-827X
  - name: Thiyanga S. Talagala
    affiliation: University of Sri Jayewardenepura
    address: Department of Statistics, Gangodawila, Nugegoda 10100 Sri Lanka
    url: https://thiyanga.netlify.app/
    email: ttalagala@sjp.ac.lk
    orcid: 0000-0002-0656-9789
type: package
creative_commons: CC BY
date: "`r Sys.Date()`"
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
  \usepackage{float}
output:
  rjtools::rjournal_article:
    keep_md: true
bibliography: paper-cardinalR.bib
editor_options:
  chunk_output_type: inline
---

### 2025-11-25 paper-cardinalR.tex
Try to find why rjtolls have the error

\abstract{%
High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association. It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold. Simulated high-dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning. These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets. By controlling the complexity researchers can better understand how different methods work and can be improved. A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets. This enriches the existing toolset of benchmark datasets for evaluating algorithms.
}

>>>>

\abstract{%
High-dimensional data is defined as each observation being described by many features, variables or dimensions, typically having some dependency or association. It is analyzed in many disciplines including image and speech recognition, clusters in genomics and forensic analysis, and sparse distributions in text mining, particularly to determine whether the observations lie on a lower dimensional manifold. Simulated high dimensional data is useful for testing, validating, and improving algorithms used in dimensionality reduction, supervised and unsupervised learning. These can be generated with a variety of structures using mathematical functions and statistical distributions to create test datasets. By controlling the complexity researchers can better understand how different methods work and can be improved. A variety of functions to generate a large variety of structures in high dimensions are organized into the R package cardinalR, along with several sample sets. This enriches the existing toolset of benchmark datasets for evaluating algorithms.
}
